---
title: "JASA Sharing Reproducible Materials over Github"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Directory structure

The repo has the following directory structure.

```{r echo = FALSE}
library(listdown)
file_type <- function() {
  a = 0
  class(a) <- "file"
  a
}
dir_type <- function() {
  a = 0
  class(a) <- "dir"
  a
}

`project directory` <- list(
  "README.md" = file_type(),
  "revn/" = dir_type(),
  "preprocessing/" = dir_type(),
  "analysis/" = dir_type(),
  "data/" = dir_type(),
  "output/" = dir_type()
)
ld_cc_dendro(`project directory`)
```

Structure of the subdirectories includes:

* `preprocessing/`: Source files for executing core modeling code, generating simulated data, and running simulations
  * `estimate_flode.R`: Defines function for estimating `flode` model using EM algorithm
  * `make_pffr.R`: Defines function for running functional historical regression model and processing surfaces for visualization
  * `search_alpha.R`: Defines function to perform grid search over possible $\alpha_0$ values to choose best initial value. 
  * `simulated_paw.R`: Defines function for simulating data.
  * `utils.R`: Defines helper functions called in the `estimate_flode.R` file.
  * `simulations/`: File for generating simulation results across all simulation scenarios.
    * `20211227_simulations_flodeVSfhist.R`: Compares flode and functional historical regression across 9 alpha values for 30 datasets each.
    * `20211228_simulations_flode.R`: Evaluates flode across 4 sample sizes and 3 alpha values for 30 datasets each.
* `analysis/`
  * `202112_dataAnalysis.Rmd/.html`: Code/output for processing raw paw trajectory data and generating Figures 1, 6, and 7 in the manuscript.
  * `202112_simulation_results.Rmd/.html`: Code/output for processing simulation results and generating Figures 2, 3, 4, and 5 in the manuscript.
  * `202112_simulation_results_cache`:




## Notes about this repo/workflow

Describe file structure and workflow here

* Slow steps are cached or saved as R data objects that are read in the next time a file is run- this makes it faster to run the Rmarkdown files that generate analysis results
* The file `202112_simulation_results.Rmd` creates and saves Figures 2, 3, 4 and 5 and stores them as `.png` files in the `output` folder
* The file `202112_dataAnalysis.Rmd` creates and saves Figures 1, 6, and 7 and stores them in the `output` folder. Also creates and saves to `output` a dataframe of computation times and parameter values that are reported as part of inline text in the `manuscript.Rmd`. This allows all inline `R` code from the `manuscript.Rmd` to  be fully reproducible (not entered by hand).
* To generate the manuscript knit the file `manuscript.Rmd` in the `manuscript` folder. This does not execute the full analysis but reads in figures and numbers in the `output` folder generated by knitting `202112_simulation_results.Rmd` and `202112_dataAnalysis.Rmd`.



# Thoughts

* easily could have this same format whether manuscripts are generated using R Markdown or Latex
* probably best to have users clone the repo rather than fork it so they can rename it?