---
title: "Data analysis figures and results"
author: "Julia Wrobel"
date: '`r format(Sys.time(), "%Y-%m-%d")`'
output:
  html_document: 
    code_folding: hide
    toc: true
    toc_float: true
hitheme: tomorrow
highlighter: highlight.js
---

# Overview

This file loads produces Figures and runs analysis using the real data.

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(stringr)
library(viridis)
library(splines)
library(refund)
library(Matrix)
library(mvtnorm)
library(patchwork)
library(gridExtra)
library(pracma)
library(tictoc)

knitr::opts_chunk$set(echo = TRUE,
  warning = FALSE,
  fig.width = 9,
  fig.height = 4,
  fig.path = '../output/'
)

theme_set(theme_bw() + theme(legend.position = "bottom"))

source(here::here("preprocessing", "estimate_flode.R"))
source(here::here("preprocessing", "utils.R"))
source(here::here("preprocessing", "search_alpha.R"))
```

# Load paw data

```{r data_import}

load(here::here("data", "20180418_tidied_data.RDA"))
```

# Data info


Get average lift and grab stats.

```{r}

paw_position %>% 
  filter(type == "Control") %>%
  summarize(mean(lift, na.rm = TRUE), sd(lift, na.rm = TRUE), 
            mean(grab, na.rm = TRUE), sd(grab, na.rm = TRUE))

```

# Data cleaning

Join the datasets

```{r}
observed_data = 
	left_join(
		seq_train %>% select(trial, type, gpfa) %>% unnest(cols = c(gpfa)),
	  paw_position %>% unnest(cols = c(paw_position))) %>% 
	mutate(sec_since_lift = seconds - lift,
	       sec_since_grab = seconds - grab) %>% 
	filter(sec_since_lift > -.05, sec_since_lift < .25,
				 type == "Control") %>%
	select(trial, sec_since_lift, px, py, pz) %>% 
	gather(key = axis, value = observed, px:pz)
```

It is necessary that the data be evaluated on the same evenly spaced grid for each subject. In order to obtain this I interpolate the data to a different grid. 

Also process raw spike trains for plotting. Also look at the paw trajectories before aligning them at the point of lift.

```{r}
D_paw = 30
make_even_grid = function(data){
	time = seq(-0.05, 0.25, length.out = D_paw)
	even_fit = approx(data$sec_since_lift, data$value, xout = time, rule = 2)$y
	
	data.frame(sec_since_lift = time, value = even_fit)
}

paw_df = 
	left_join(
		seq_train %>% select(trial, type, gpfa) %>% unnest(cols = c(gpfa)),
		paw_position %>% unnest(cols = c(paw_position))) %>% 
    mutate(sec_since_lift = seconds - lift) %>% 
    filter(sec_since_lift > -.05, sec_since_lift < .25, type == "Control") %>% 
	select(trial, type, sec_since_lift, px, py, pz, starts_with("gpfa")) 

even_paw = paw_df %>%
	select(trial, type, sec_since_lift, starts_with("p")) %>%
	gather(axis, value, starts_with("p")) %>%
	nest(data = c(sec_since_lift, value)) %>%
	mutate(even_data = map(data, make_even_grid)) %>% 
	unnest(even_data) %>%
	mutate(time = sec_since_lift - min(sec_since_lift), u = 1) %>%
	select(trial, axis, sec_since_lift, time, u, value)

even_gpfa = paw_df %>%
	select(trial, type, sec_since_lift, starts_with("g")) %>%
	gather(gpfa, value, starts_with("g")) %>%
	nest(data = c(sec_since_lift, value)) %>%
	mutate(even_data = map(data, make_even_grid)) %>% 
	unnest(even_data) %>%
	mutate(time = sec_since_lift - min(sec_since_lift)) %>%
	rename(u_gpfa = value) %>%
	select(trial, time, gpfa, u_gpfa) %>%
	spread(gpfa, u_gpfa)
	

paw_df = left_join(even_paw, even_gpfa)

firing_df = 
	left_join(
		seq_train %>% select(trial, type, neuron_fire) %>% 
		  filter(type == "Control") %>%
		  mutate(row = row_number()) %>% unnest(cols = c(neuron_fire)),
		paw_position %>% filter(type == "Control") ) %>% 
    #mutate(sec_since_lift = seconds - lift) %>% 
    #filter(sec_since_lift > -.05, sec_since_lift < .25) %>% 
  filter(seconds < 2) %>%
  #mutate(time = sec_since_lift - min(sec_since_lift)) %>%
	#select(row, trial, type, sec_since_lift, time, seconds, starts_with("loc")) 
  arrange(type) %>%
  select(row, trial, seconds, starts_with("loc"))

```


# Analysis


## data organization

We run the model on the three axes separately. 


```{r}
paw_x = paw_df %>%
  filter(axis == "px") %>%
  group_by(trial) %>%
  mutate(int = 1,
         y0 = first(value),
         time = time/0.3)  %>%
  ungroup()

paw_y = paw_df %>%
  filter(axis == "py") %>%
  group_by(trial) %>%
  mutate(int = 1,
         y0 = first(value),
         time = time/0.3)  %>%
  ungroup()

paw_z = paw_df %>%
  filter(axis == "pz") %>%
  group_by(trial) %>%
  mutate(int = 1,
         y0 = first(value),
         time = time/0.3)  %>%
  ungroup()

```



## modeling

First we do a grid search to find the best initial value of $\alpha$. For the grid search we run the model without random effects, choosing a particular value of $\alpha$ and using the observed initial positions to calculate the value of the loss function at that particular $\alpha$.

Run flode on the data and save flode data object. Only run the flode analysis if saved object doesn't exist.


### px


```{r px}
filename = here::here("output", "px_results.RData")

if(file.exists(filename)){
	load(filename)
}else{
	tic()
	# initiate alpha
	alpha_start_px = search_alpha(data = paw_x, 
														 forcing_functions = c("int", "gpfa_1", "gpfa_2", "gpfa_3", 
																									"gpfa_4", "gpfa_5"),
														 Kt = 20)
	
	# run flode
	px_results = estimate_flode(paw_x, 
															alpha0 = alpha_start_px$best_alpha, 
                            	Kt = 20,
															forcing_functions = c("int", "gpfa_1", "gpfa_2", "gpfa_3", 
																										"gpfa_4", "gpfa_5"), 
															max_iter = 500, 
															tol = 0.000001, 
															initial_position = TRUE,
															random_effects = TRUE)
	time_px = toc()

	save(px_results, time_px, alpha_start_px, file = filename)
}


```

### py


```{r py}
filename = here::here("output", "py_results.RData")

if(file.exists(filename)){
	load(filename)
}else{
	tic()
	# initiate alpha
	alpha_start_py = search_alpha(data = paw_y, 
														 forcing_functions = c("int", "gpfa_1", "gpfa_2", "gpfa_3", 
																									"gpfa_4", "gpfa_5"),
														 Kt = 20)
	
	# run flode
	py_results = estimate_flode(paw_y, 
															alpha0 = alpha_start_py$best_alpha, 
                            	Kt = 20,
															forcing_functions = c("int", "gpfa_1", "gpfa_2", "gpfa_3", 
																										"gpfa_4", "gpfa_5"), 
															max_iter = 500, 
															tol = 0.000001, 
															initial_position = TRUE,
															random_effects = TRUE)
	time_py = toc()

	save(py_results, time_py, alpha_start_py, file = filename)
}

```

### pz


```{r pz}
filename = here::here("output", "pz_results.RData")

if(file.exists(filename)){
	load(filename)
}else{
	tic()
	# initiate alpha
	alpha_start_pz = search_alpha(data = paw_z, 
														 forcing_functions = c("int", "gpfa_1", "gpfa_2", "gpfa_3", 
																									"gpfa_4", "gpfa_5"),
														 Kt = 20)
	
	# run flode
	pz_results = estimate_flode(paw_z, 
															alpha0 = alpha_start_pz$best_alpha, 
                            	Kt = 20,
															forcing_functions = c("int", "gpfa_1", "gpfa_2", "gpfa_3", 
																										"gpfa_4", "gpfa_5"), 
															max_iter = 500, 
															tol = 0.000001, 
															initial_position = TRUE,
															random_effects = TRUE)
	time_pz = toc()

	save(pz_results, time_pz, alpha_start_pz, file = filename)
}

```


# Summaries

Below are the computation times for each axis.

```{r}
summary_stats = tibble(time_px = round(as.numeric( (time_px$toc-time_px$tic)/60), 2),
       time_py = round(as.numeric( (time_py$toc-time_py$tic)/60), 2),
       time_pz = round(as.numeric( (time_pz$toc-time_pz$tic)/60), 2),
       alpha_px = px_results$alpha,
       alpha_py = py_results$alpha,
       alpha_pz = pz_results$alpha)

# save summary stats so they can be read into manuscript Rmd file
save(summary_stats, file = "../output/summarystats.rda")
```



* px: `r summary_stats$time_px` minutes
* py: `r summary_stats$time_py` minutes
* pz: `r summary_stats$time_pz` minutes


Below are the estimated alpha for each axis:

* px: $\alpha =$ `r summary_stats$alpha_px`
* py: $\alpha =$ `r summary_stats$alpha_py`
* pz: $\alpha =$ `r summary_stats$alpha_pz`


# Figures

## Figure 1

Below are plots of the paw trajectories and gpfa values.

* Called `fig_data_eda` in the manuscript

```{r fig_data_eda, fig.width = 10, fig.height= 8}
paw_plot = paw_df %>%
  mutate(axis = case_when(
    axis == "px" ~ "x-axis",
    axis == "py" ~ "y-axis",
    axis == "pz" ~ "z-axis"
  )) %>%
	ggplot(aes(sec_since_lift, value, group = trial)) +
	geom_line(alpha = 0.25) +
	labs(x = "seconds since lift", y = "position") +
	facet_wrap(~axis, scales = "free_y") #+
  #theme(axis.title.x = element_blank())


neuron_plot = firing_df %>%
    select(row, trial, seconds, loc_02, loc_06, loc_09) %>%
    gather(key = loc, value = value, starts_with("loc")) %>%
    mutate(spike = as.numeric( value > 0),
           loc = str_replace(loc, "loc_0", "location ")) %>%
    ggplot(aes(x = seconds, y = row)) + 
    geom_tile(aes(fill = factor(spike))) +
    scale_fill_viridis(discrete = TRUE) + 
    #scale_fill_manual(values = c("lightblue", "darkblue")) +
		labs(y = "trial") + 
		facet_grid(.~loc) + 
    theme(legend.position = "none",
          axis.ticks=element_blank(),
          axis.line=element_blank(),
          axis.text.x = element_blank(),
          axis.title.x = element_blank()) +
    scale_x_continuous(expand=c(0,0)) + 
    scale_y_continuous(expand=c(0,0)) 


gpfa_plot = paw_df %>% 
	filter(axis == "px") %>%
	gather(gpfa, x_value, gpfa_1:gpfa_5) %>%
  mutate(gpfa = str_replace(gpfa, "_", " ")) %>%
	ggplot(aes(sec_since_lift, x_value, group = trial)) +
	geom_line(alpha = 0.2) +
	labs(x = "seconds since lift", y = "value") +
	facet_wrap(~ gpfa, scales = "free_y", ncol = 5)


#p1 + (p2 + p3) + p4 + plot_layout(ncol = 1)
(paw_plot + neuron_plot) + gpfa_plot + plot_layout(ncol = 1 )
```

## Figure 6

* Called `fig_data_fits` in the manuscript

### Data org

Combine datasets from the three axes. 

```{r}
x_df = px_results$data %>% 
  select(trial, time, value, yhat, delta, deltaStar) %>%
  mutate(axis = "px") 

y_df = py_results$data %>% 
  select(trial, time, value, yhat, delta, deltaStar) %>%
  mutate(axis = "py")

z_df = pz_results$data %>% 
  select(trial, time, value, yhat, delta, deltaStar) %>%
  mutate(axis = "pz")


flode_results = bind_rows(x_df, y_df, z_df) %>%
  mutate(time = time * 0.3)

```


### Plots

```{r fig_data_fits, fig.height= 10}
fits = flode_results %>%
	  mutate(axis = case_when(
    axis == "px" ~ "x-axis",
    axis == "py" ~ "y-axis",
    axis == "pz" ~ "z-axis"
  )) %>%
  ggplot(aes(time, value, group = trial)) + 
  geom_line(alpha = 0.5, color = "gray") +
  geom_line(aes(y = yhat), alpha = 0.1, color = "darkred") +
  geom_vline(xintercept = 0.05, linetype = 2, color = "blue") +
	labs(y = "Y(t)") +
  facet_wrap(~ axis, scales = "free")

delta = flode_results %>%
		  mutate(axis = case_when(
    axis == "px" ~ "x-axis",
    axis == "py" ~ "y-axis",
    axis == "pz" ~ "z-axis"
  )) %>%
  ggplot(aes(time, delta, group = trial)) + 
  geom_line(alpha = 0.1) +
  geom_vline(xintercept = 0.05, linetype = 2, color = "blue") +
	labs(y = "random effects, derivative scale") +
  facet_wrap(~ axis)

deltaStar = flode_results %>%
		  mutate(axis = case_when(
    axis == "px" ~ "x-axis",
    axis == "py" ~ "y-axis",
    axis == "pz" ~ "z-axis"
  )) %>%
  ggplot(aes(time, deltaStar, group = trial)) + 
  geom_line(alpha = 0.1) +
  geom_vline(xintercept = 0.05, linetype = 2, color = "blue") +
	labs(y = "random effects, data scale") +
  facet_wrap(~ axis)

fits  +  deltaStar + delta  + plot_layout(ncol = 1)
```


## Figure 7

* Called `fig_data_beta` in the manuscript



### Data org


```{r}
# betas
beta_results = bind_rows(px_results$beta %>% mutate(axis = "px"),
          py_results$beta %>% mutate(axis = "py"),
          pz_results$beta %>% mutate(axis = "pz")) %>%
  mutate(time = time * 0.3)

```

### Plot


```{r fig_data_beta}
beta_results %>%
  gather(beta, value, starts_with("beta"))  %>%
  ggplot(aes(time, value, group = axis, color = axis)) + 
	geom_line() +
  geom_vline(xintercept = 0.05, linetype = 2, color = "black") +
  facet_wrap(~ beta) +
  theme(legend.position = "bottom")

```


## Figure 8

* Called `fig_data_surfaces` in the manuscript


### Data org

```{r}

## for surfaces just look at paw X
# now get surfaces
make_surface_df = function(data, beta_num){
  surface = data$surface[[beta_num]]
  
  
  surface %>% as_tibble() %>%
    mutate(t = attributes(surface)$t) %>%
    gather(s, value, starts_with("s")) %>%
    arrange(t) %>%
    mutate(s = rep(attributes(surface)$s, 30),
           label = paste0("beta", beta_num - 1))
}

surface_df = map_dfr(2:6, make_surface_df, data = px_results)

```

### Plot


```{r fig_data_surfaces}
surface_df %>%
  mutate(value = ifelse(s <= t, value, NA),
         s = s * 0.3,
         t = t * 0.3) %>%
  ggplot(aes(s, t)) + 
  geom_tile(aes(fill = value, col = value)) + 
  scale_fill_viridis_c() +
  scale_colour_viridis_c() +
  facet_wrap(~ label, , ncol = 5) +
  theme(legend.position = "bottom")


```
